Towards a free/open-source universal-dependency treebank for Kazakh

### 2015-06-21 2015-turklang-kazdep/paper.tex
commit
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60936 72bbbca6-d526-0410-a7d9-f06f51895060

Non-finite

@@

\begin{table}[htbp]
  \centering
  \caption{Composition of the corpus. The corpus covers a range of genres and text types from free and public-domain sources.}
	\begin{small}
		\begin{tabular}{llrrr}
			\toprule
				\textbf{Document} & \textbf{Description} & \textbf{Sentences} & \textbf{Tokens} & \textbf{Avg. length}\\
			\midrule
				UN Declaration on Human Rights & Legal text on human rights & 25 & 409 & 16.3 \\
				Phrasebook                     & Phrases from Wikitravel   & 37 & 205 & 5.5 \\
				Жиырма Бесінші Сөз             & Philosophical text        & 34 & 525 & 15.4 \\ 
				Қожанасырдың тойға баруы       & Folk tale from Wikisource  & 8 & 140 & 17.5 \\
				Ер Төстік                      & Folk tale from Wikisource  & 23 & 203 & 8.8 \\
				Азамат қайда?                  & Story for language learners & 48 & 434 & 9.0 \\
				Футболдан әлем чемпионаты 2014 & Wikipedia article (2014 World Cup) & 14 & 246 & 17.5 \\
				Иран & Wikipedia article (Iran)                                & 111 & 1562 & 14.0 \\
				Радиан & Wikipedia article (Radian)                            & 2 & 17 & 8.5 \\
			\midrule
			\multicolumn{2}{c}{\tilda{}} & 302 & 3741 & 12.3 \\\cmidrule[\heavyrulewidth]{3-5}
		\end{tabular}
	\end{small}
\end{table}

>>>>

Kazakh, like most Turkic languages, makes frequent use of non-finite verb forms by deriving verbal adjectives, verbal nouns, and verbal adverbs from verbs.  Verbal adjective phrases modify nouns, as in \kazakh{мені \textbf{көрген} дәрігер} \gloss{the doctor that saw me}.  Verbal nouns can function as subjects or complements of verbs or copulas, as in \kazakh{Дәрігер сені \textbf{көргенін} білген жоқпын.} \gloss{I didn't know that the doctor had seen you.} and \kazakh{Дәрігер сені \textbf{көргені} жақсы болыпты.} \gloss{It's good that doctor saw you.}

@@

\begin{table}[htbp]
  \centering
  \caption{Composition of the corpus. The corpus covers a range of genres and text types from free and public-domain sources.}
	\begin{small}
		\begin{tabular}{llrrr}
			\toprule
				\textbf{Document} & \textbf{Description} & \textbf{Sentences} & \textbf{Tokens} & \textbf{Avg. length}\\
			\midrule
				UN Declaration on Human Rights & Legal text on human rights & 25 & 409 & 16.3 \\
				Phrasebook                     & Phrases from Wikitravel   & 37 & 205 & 5.5 \\
				Жиырма Бесінші Сөз             & Philosophical text        & 34 & 525 & 15.4 \\ 
				Қожанасырдың тойға баруы       & Folk tale from Wikisource  & 8 & 140 & 17.5 \\
				Ер Төстік                      & Folk tale from Wikisource  & 23 & 203 & 8.8 \\
				Азамат қайда?                  & Story for language learners & 48 & 434 & 9.0 \\
				Футболдан әлем чемпионаты 2014 & Wikipedia article (2014 World Cup) & 14 & 246 & 17.5 \\
				Иран & Wikipedia article (Iran)                                & 111 & 1562 & 14.0 \\
				Радиан & Wikipedia article (Radian)                            & 2 & 17 & 8.5 \\
			\midrule
			\multicolumn{2}{c}{~} & 302 & 3741 & 12.3 \\\cmidrule[\heavyrulewidth]{3-5}
		\end{tabular}
	\end{small}
\end{table}
